Verifying Spatial Properties of Stencil Computations

### 2016-11-16 paper.tex
edits to last section

Various deductive verification tools can express array indexing
in their specifications, \eg{}, ACSL of \citet{baudin2008acsl}
for C (see \eg{}~\citet[Example 3.4.1]{burghardt2010acsl}). Thus, a specification
can be given for a stencil computation, 
however, such a specification would exactly replicate the fine-grained
indexing seen in the code and would simply be prone to low-level input
errors. Our approach is much more abstract-- it does
not aim to reify indexing in the specification, but is novel
in that it instead
provides abstract spatial descriptions which capture a large number
of common patterns.

\citet{kamil2016verified} propose \emph{verified lifting} to extract a
functionally-complete mathematical representation of low-level, and
potentially optimised, stencils in Fortran code. This extracted
predicate representation of a stencil is used to generate code
for the \textsc{Halide} high-performance
compiler~\citep{ragan2013halide}. Thus their approach must capture the
full meaning of a stencil computation which requires significant
program analysis. For example, they report that some degenerate
stencil kernels take up to 17 hours to analyse and others require
programmer intervention for correct invariants to be inferred.

@@

Our approach differs significantly. Rather than 
full representation of stencils, we focus on specifying
just the spatial behaviour in a lightweight way.
Thus, it suffices for us to perform a comparatively
simple data-flow analysis which is efficient, scales linearly with
code size, and does not require any user intervention.
Whilst we do not perform deep semantic analysis of stencils,
 the analysis part of our
approach can be made arbitrarily more sophisticated independent of the rest of
the work.
% Hence, we do not require SMT solving or search of loop invariants. It
%suffices to do comparatively simple data-flow analysis.
%Since CamFort mostly does
%syntactic analysis, the inference procedure terminates quickly and
%never requires programmer intervention. 
%The downside of our
%approach is that if optimisation causes the stencil computation to be heavily
%obfuscated, \textsf{STNG} would capture the access behaviour better.
%
Furthermore, Kamil \emph{et al.} do not provide a user-visible
syntactic representation of their specifications, and they do
not use their specifications for verification, \eg{}, to future-proof
the code against later changes. Even if they were to provide a
syntactic representation, for complex stencils such as Navier-Stokes
from \cref{sec:intro}, it would be as verbose as the code itself,
making it difficult for programmer to understand the overall shape of
the indexing behaviour.

@@

By contrast, the work of \citet{Abe:2013:IPDPSW} has looked at the
correctness problem explicitly by bringing a form of model-checking to
verify certain stencil computations. This is in the context of
decomposing stencils for parallel computation on multiple processors
 in partitioned global address
space languages. %These are scenarios where an array is divided into
%subarrays on multiple processors but each has the ability to access
%the others' memory. The authors avoid the state explosion dilemma of
%model-checking by relying upon the fact that accesses to the
%``boundary elements'' will require a different method by virtue of the
%non-locality of that memory. Since this access is performed in a
%different manner, those can be identified automatically by static
%analysis. To achieve this, 
Abe \emph{et al.} provide a new language to
be used for writing stencil computations. Much of the specification
effort goes towards describing the nuts and bolts of distributing the
computation over multiple processors. The code for the stencil kernel
is generated from a relatively high-level specification. 
In contrast, we are interested in integrating directly
into existing, legacy codebases and established languages, bringing
the benefits of verification more easily to scientific computing.%such as
%older versions of Fortran.% We infer stencil specifications from
%Fortran code and check annotations on stencil computations within
%Fortran code.

>>>>

Various deductive verification tools can express array indexing
in their specifications, \eg{}, ACSL of \citet{baudin2008acsl}
for C (see \eg{}~\citet[Example 3.4.1]{burghardt2010acsl}). Thus, a specification
can be given for a stencil computation, 
however, such a specification would exactly replicate the fine-grained
indexing seen in the code and would be similarly prone to low-level input
errors. Our approach is much more abstract-- it does
not aim to reify indexing in the specification, but is novel
in that it instead
provides abstract spatial descriptions which capture a large number
of common patterns.

\citet{kamil2016verified} propose \emph{verified lifting} to extract a
functionally-complete mathematical representation of low-level, and
potentially optimised, stencils in Fortran code. This extracted
predicate representation of a stencil is used to generate code
for the \textsc{Halide} high-performance
compiler~\citep{ragan2013halide}. Thus they must capture the
full meaning of a stencil computation which requires significant
program analysis. For example, they report that some degenerate
stencil kernels take up to 17 hours to analyse and others require
programmer intervention for correct invariants to be inferred.

@@

Our approach differs significantly. Rather than 
full representation of stencils, we focus on specifying
just the spatial behaviour in a lightweight way.
Thus, it suffices for us to perform a comparatively
simple data-flow analysis which is efficient, scales linearly with
code size, and does not require any user intervention.
Whilst we do not perform deep semantic analysis of stencils,
 the analysis part of our
approach can be made arbitrarily more sophisticated independent of the rest of
the work.
% Hence, we do not require SMT solving or search of loop invariants. It
%suffices to do comparatively simple data-flow analysis.
%Since CamFort mostly does
%syntactic analysis, the inference procedure terminates quickly and
%never requires programmer intervention. 
%The downside of our
%approach is that if optimisation causes the stencil computation to be heavily
%obfuscated, \textsf{STNG} would capture the access behaviour better.
%
Furthermore, Kamil \emph{et al.} do not provide a user-visible
syntactic representation of their specifications, and nor do
they provide verification from specifications \eg{}, to future-proof
the code against later changes. Even if they were to provide a
syntactic representation, for complex stencils such as Navier-Stokes
from \cref{sec:intro}, it would be as verbose as the code itself,
making it difficult for programmer to understand the overall shape of
the indexing behaviour.

@@

By contrast, the work of \citet{Abe:2013:IPDPSW} has looked at the
correctness problem explicitly by bringing a form of model-checking to
verify certain stencil computations. This is in the context of
decomposing stencils for parallel computation on multiple processors
 in partitioned global address
space languages. %These are scenarios where an array is divided into
%subarrays on multiple processors but each has the ability to access
%the others' memory. The authors avoid the state explosion dilemma of
%model-checking by relying upon the fact that accesses to the
%``boundary elements'' will require a different method by virtue of the
%non-locality of that memory. Since this access is performed in a
%different manner, those can be identified automatically by static
%analysis. To achieve this, 
Abe \emph{et al.} provide a new language for writing stencil computations. Much of the specification
effort goes towards describing the distribution of the
computation over multiple processors. The code for the stencil kernel
is generated from a relatively high-level specification. 
In contrast, we are interested in integrating directly
into existing, legacy codebases and established languages, bringing
the benefits of verification more easily to scientific computing.%such as
%older versions of Fortran.% We infer stencil specifications from
%Fortran code and check annotations on stencil computations within
%Fortran code.
